---
title: "Reproducible Research Project 1"
author: "Tessa Jones"
date: "January 20, 2016"
output: html_document
---

The following outlines the basic analysis of data retrieved from a personal activity monitoring device that measures the number of steps taken by an individual at 5 m inute intervals over the course of two months. 

Step one is to import to csv file with the activity data.

```{r, echo=TRUE}
library(plyr)
setwd('C:/excercisedata/repdata-data-activity')
data <- read.csv('activity.csv', stringsAsFactors= FALSE)
```

The total number of steps taken per day is summarized below and displayed as a histogram.

```{r, echo=TRUE}
sumbyday <- ddply(data, .(date), summarize, sum = sum(steps, na.rm = TRUE))
hist(sumbyday$sum, xlab = 'Days', ylab = 'Steps', main = 'Steps per Day')
```

The mean number of steps taken per day is 37.3826 and the median is 0

```{r, echo=TRUE}
mean(data$steps, na.rm = TRUE)
median(data$steps, na.rm = TRUE)
```

A time series plot below was generated to identify the average daily activity.  The 5 minute time interval that, on average, has the most number of steps is 835.  

```{r, echo=TRUE}
meanbyint <- ddply(data, .(interval), summarize, mean=mean(steps, na.rm = TRUE))
library(ggplot2)
ggplot(meanbyint, aes(interval, mean)) + geom_line() + xlab("5 Minute Interval of Day") + ylab("Number of Steps")
```

There are 2304 missing values from the data set, here I have replaced the missing values with the average number of steps for that given time interval. The following produces a histrogram of the sum of steps for each day using the new dataset.  There is small differences between the two histograms (before and after accounting for the missing values). The first few weeks have a lower number of steps per day afterwards.  The mean and median values remained the same both before and after the data replacement.
```{r, echo=TRUE}
sum(is.na(data$steps))
sub1 <- merge(x = data, y = meanbyint, by = "interval", all.y = TRUE)
sub2 <- transform(sub1, steps=ifelse(is.na(sub1$steps),sub1$mean, sub1$steps))
sumbyday2 <- ddply(sub2, .(date), summarize, sum = sum(steps, na.rm = TRUE))
hist(sumbyday2$sum, xlab = 'Days', ylab = 'Steps', main = 'Steps per Day')
mean(sub2$steps, na.rm = TRUE)
median(sub2$steps, na.rm = TRUE)
```

Now we want to compare the differences in activity between weekends and weekdays.  The dates have been defined as being either a weeday or a weekend and a panel plot was generated to differentiate between the steps at each interval between weekends and weekdays.  
```{r, echo=TRUE}
sub2$day <- weekdays(as.Date(sub2$date))
sub2[sub2$day %in% c("Saturday", "Sunday"),]$day <- "Weekend"
sub2[sub2$day %in% c("Monday", "Tuesday","Wednesday","Thursday","Friday"),]$day <- "Weekday"
sub2 <- ddply(sub2, .(interval,day), summarise, meanReal=mean(steps))
library(lattice)
xyplot(sub2$meanReal~sub2$interval|sub2$day, type = 'l', xlab = 'Interval', ylab = 'Mean')
```


