---
title: "Coursera Project 1"
author: "Tessa Jones"
date: "Friday, August 14, 2015"
output: html_document
---

This is the results for Project 1 of the Reproducible Research Course for Coursera:

```{r}
setwd("C:/Users/tejones/Desktop/Coursera/Knitr")
data <- data.frame(read.csv("activity.csv"))
data2 <- na.omit(data)
library(plyr)


dailysum <- ddply(data2, .(date),summarise, sum=sum(steps))
```
Total number of steps taken per day
```{r}
dailysum

```
Histogram of the sum of steps taken per day is shown below
```{r}
hist(dailysum$sum, xlab = 'Days', ylab = 'Steps', main = 'Steps per Day')
```


Mean of the Steps taken per day is 10766.19 steps 
```{r}
mean(dailysum$sum)
```
Median of the Steps taken per day is 10765 steps
```{r}
median(dailysum$sum)

dailyinterval <- ddply(data2, .(interval),summarise, mean=mean(steps))

plot(dailyinterval$interval, dailyinterval$mean, type = 'l', xlab = 'Interval', ylab = 'Average Steps', main = 'Average Steps Per Interval')

```


The interval with the maximum mean is 835, as shown below
```{r}
dailyinterval[dailyinterval$mean == max(dailyinterval$mean),]

```
The number of NA values is 2304, as shown below
```{r}
sum(is.na(data))

intervalmerge <- merge(data, dailyinterval, by.x = 'interval', by.y = 'interval')

intervalmerge$steps[is.na(intervalmerge$steps)] <- intervalmerge$mean[is.na(intervalmerge$steps)]

dailysum2 <- ddply(intervalmerge, .(date),summarise, sum=sum(steps))

hist(dailysum2$sum, xlab = 'Days', ylab = 'Steps', main = 'New Steps per Day')
```


The NA values were replaced with the total mean interval value and produced the results above
```{r}
 
intervalmerge$day <- weekdays(as.Date(dailysum2$date))

intervalmerge[intervalmerge$day %in% c("Saturday","Sunday"),]$day <- "Weekend"

intervalmerge[intervalmerge$day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),]$day <- "Weekday"

intervalmerge <- ddply(intervalmerge, .(interval,day),summarise, mean=mean(steps))

library(lattice)

xyplot(intervalmerge$mean~intervalmerge$interval|intervalmerge$day, type = 'l', xlab = 'Interval', ylab = 'Mean')

```


Above is a representation of mean interval values for Weedays vs. Weekends
```{r}
```


